---
title: "Notes for 2-12"
format: 
  html:
    self-contained: true
# format: revealjs
df-print: tibble
knitr:
    opts_knit: 
      root.dir: "~/marginalization"
    opts_chunk: 
      echo: true
      output: true
editor_options: 
  chunk_output_type: console
---

## Highlights

* The Stan model is sensitive to the prior proportion of nulls ($\pi_0$)
* The Stan model with Poisson response has a very poor fit
  * Offsetting biased estimates of several parameters
* The 'Ribodiff' model allows different dispersions for RNA and RPF samples
in riboseq

### $\pi_0 = 0.8$

I found an error - I was setting $\pi_0 = 0.2$ ($\pi_0$ is the prior proportion of null genes).
The true proportion of null genes is $0.8$. Setting $\pi_0 = 0.8$ instead of 
to an incorrect value improves the performance of the stan models. 

```{r}
#| echo: false
readRDS("cs_BYN1/roc_plot_8.rds")
readRDS("cs_BYN1/fdr_plot_8.rds")
readRDS("cs_BYN1/b1_plot_8.rds")
# readRDS("cs_BYN1/phi_plot_8.rds")
```

### Poisson Stan fit

The sample normfactors plot isn't great because of the big difference in y-axis scale, but 
the Poisson response model is completely missing (overestimating by a large margin) the true value,
where the NB response model's posterior is centered right at the true value. 

```{r}
#| echo: false
readRDS("cs_BYN1/b0_plot_9.rds")
readRDS("cs_BYN1/S_plot_9.rds")
```

### Ribodiff

- We saw quite poor performance from this method in the ribovi paper (and I think also in one of Ran's papers?)
- Their model does allow different dispersions for RNA and RPF samples
- Their model does not have any pairing between RNA and RPF from the same biological sample

## Extra


### Adjusting the prior proportion of nulls

I ran the model with
a few different settings for the $\pi_0$ prior to test sensitivity. 

#### $\pi_0 = 0.9$

```{r}
#| echo: false
readRDS("cs_BYN1/roc_plot_9.rds")
readRDS("cs_BYN1/fdr_plot_9.rds")
readRDS("cs_BYN1/b1_plot_9.rds")
# readRDS("cs_BYN1/phi_plot_9.rds")
```

#### $\pi_0 = 0.99$

```{r}
#| echo: false
readRDS("cs_BYN1/roc_plot_99.rds")
readRDS("cs_BYN1/fdr_plot_99.rds")
readRDS("cs_BYN1/b1_plot_99.rds")
# readRDS("cs_BYN1/phi_plot_99.rds")
```

### Simulation description

I replicated a simple simulation setting from [BYN PLOSOne 2020](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0232271#pone.0232271.ref021)
and compared the performance of edgeR and my Stan models on this data set. 
The setting is just a two-group comparison, with baseline means and dispersions
sampled from edgeR estimates produced by a large real data set and FC simulated from 1.5 + exp(1). 
There are no random effects. 

- Stan model with negative binomial response and estimated dispersion is ok
  - Better logFC estimates than edgeR
  - Similar ROC to edgeR
  - Liberal FDR control, where edgeR is conservative
  - Overestimates (truly) large dispersions, where edgeR underestimates 
- Stan model with Poisson response is quite poor