---
title: "Notes for 2-26"
format: 
  html:
    self-contained: true
# format: revealjs
df-print: tibble
knitr:
    opts_knit: 
      root.dir: "~/marginalization"
    opts_chunk: 
      echo: false
      output: true
editor_options: 
  chunk_output_type: console
---

## Conditionally Poisson? 

- Does it make sense to use a Poisson model for counts conditional on random effects? 
- Impression of the literature: 
  - Early identification of overdispersion
  - Popular methods use NB
  - NB is more flexible, so doesn't need much justification
- How would we check if Poisson was adequate after conditioning on random effects? 
- Naive idea: use AIC to compare NB and Poisson GLM

### Simple simulation

- Two treatments
- Paired samples

Mean model for each count: 

$$
log(\mu_{gi}) = x_i^T\beta_g + z_i^T u_g + log(S_i)
$$
Then $y_{gi} \sim NB(\mu_{gi}, \phi_g)$ or $y_{gi} \sim Pois(\mu_{gi})$; 
each simulated data set uses either NB or Pois. 

Naive procedure: 

- Estimate the $S_i$ before model fitting and treat as offsets
- Fit either Poisson GLMs or NB GLMs to each row 
- Compare AIC between the two fits for each row

Fit GLMs using this formula: 

```{r}
#| eval: false
#| echo: true
formula(y ~ trt + pairs + offset(log_S))
```

Preferred model when counts are truly NB distributed, conditional on pair effects (out of 100): 

```{r}
w_pairs <- readRDS("aic_w_pairs.rds")
summary(w_pairs$aic_nb$preferred_model)
```

Preferred model when counts are truly Poisson distributed, conditional on pair effects (out of 100): 

```{r}
summary(w_pairs$aic_pois$preferred_model)
```

Aside: if the model doesn't include the pair effects, in other words if 
the model is fit with this formula: 

```{r}
#| eval: false
#| echo: true
formula(y ~ trt + offset(log_S))
```

Then the NB model is generally preferred no matter the true conditional response
distribution (which I think makes sense). 

### KIRC data

The KIRC kidney data set includes 36 pairs of samples (tumor and non-tumor). 
I repeated the procedure described above on 1000 genes from this data set. 
The negative binomial GLM  has lower AIC for all 1000 of these genes.

```{r}
readRDS("kirc_aic_plot.rds")
```

## Extra

Papers that are referenced for overdispersion (not a complete list - I quit grabbing links for these):  

- [Baggerly 1](https://pmc.ncbi.nlm.nih.gov/articles/PMC524524/)
- [Lu](https://pmc.ncbi.nlm.nih.gov/articles/PMC1189357/)