---
title: "Simple Published Simulation Replication + Zika Dispersion Comparisons"
format: 
  html:
    self-contained: true
# format: revealjs
df-print: tibble
knitr:
    opts_knit: 
      root.dir: "~/marginalization"
    opts_chunk: 
      echo: true
      output: true
editor_options: 
  chunk_output_type: console
---
 
## Discussion

### Stan models

I replicated a simple simulation setting from [BYN PLOSOne 2020](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0232271#pone.0232271.ref021)
and compared the performance of edgeR and my Stan models on this data set. 
The setting is just a two-group comparison, with baseline means and dispersions
sampled from edgeR estimates produced by a large real data set and FC simulated from 1.5 + exp(1). 
There are no random effects. 

- Stan model with negative binomial response and estimated dispersion is ok
  - Better logFC estimates than edgeR
  - Similar ROC to edgeR
  - Liberal FDR control, where edgeR is conservative
  - Overestimates (truly) large dispersions, where edgeR underestimates 
- Stan model with Poisson response is quite poor

### Zika ribo-seq data set

I compared dispersion estimates produced by DESeq2 for RNA and RPF samples separately. 
The estimates do not appear strongly correlated. 

## Stan Models

```{r}
#| echo: false
readRDS("cs_BYN1/roc_plot.rds")
readRDS("cs_BYN1/fdr_plot.rds")
readRDS("cs_BYN1/b1_plot.rds")
readRDS("cs_BYN1/phi_plot.rds")
```

## Zika ribo-seq 

### Dispersions for RNA samples vs. RPF samples

```{r}
#| echo: false
readRDS("zika_res/JEG3_WT_disp_plot.rds")
readRDS("zika_res/JEG3_MUTII_disp_plot.rds")
readRDS("zika_res/WT_MUTII_disp_plot.rds")
```

### logFC estimates

```{r}
#| echo: false
readRDS("zika_res/JEG3_WT_type_lfc_rna_plot.rds")
readRDS("zika_res/JEG3_MUTII_type_lfc_rna_plot.rds")
readRDS("zika_res/WT_MUTII_type_lfc_rna_plot.rds")
readRDS("zika_res/JEG3_WT_type_lfc_rpf_plot.rds")
readRDS("zika_res/JEG3_MUTII_type_lfc_rpf_plot.rds")
readRDS("zika_res/WT_MUTII_type_lfc_rpf_plot.rds")
```